---
title: "hb-data-vis"
author: "Sarah Supp"
date: "October 4, 2016"
output: html_document
params:
  species:
    label: "Species:"
    value: ruhu
    input: select
    choices: [bchu, bthu, cahu, rthu, ruhu]
  dropbox: 
    label: "Input_dropbox_path:"
    value: /home/sarah/Dropbox/Hummingbirds/
    input: text
  github:
    label: "Input_github_path"
    value: /home/sarah/Documents/GitHub/hb-migration/
    input: text
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

## Getting set up:
Note: pmin is backwards (all previous points) and ppls is forwards (all future points)

```{r, echo=FALSE}
# to run, type:
#    rmarkdown::render("/home/sarah/Documents/GitHub/hb-migration/hb-data-vis.Rmd", params = "ask")
library(dplyr)
library(ggplot2)
library(ggmap)
library(gganimate)
library(lubridate)
library(plyr)
library(plotly)
library(gridExtra)
library(viridis)
library(ggthemes)

github <- params$github
dropbox <- params$dropbox

#assign based on the time frame (number of days) used to compute the alpha hulls
alpha_window = 5

# define pathnames (FAL will need to change pathnames)
function.dir <- paste0(github, "hb_RS_functions.R")
agan.dir <- paste0(dropbox, "hb_migration_data/ebird_annotated_raw/combined/")
abs.dir <- paste0(dropbox, "hb_migration_data/pseudoabsences/t", alpha_window, "/annotated/")
migtime.dir <- paste0(dropbox, "hb_migration_data/ebird_raw/eBird_checklists_2008-2014/aggregate_by_species/")
fig.dir <- paste0(dropbox, "NASA_Hummingbirds/P10_eBird_Migration_multiple topics/2-Mechanisms/figures/")
stat.dir <- paste0(dropbox, "NASA_Hummingbirds/P10_eBird_Migration_multiple topics/2-Mechanisms/stat-tests/")

source(function.dir)

sp = params$species
```

# `r sp`
## Import data to generate species report
```{r, echo=FALSE}

  spfiles = list.files(path = agan.dir, pattern = glob2rx(paste0(sp,"*.RData")), recursive=FALSE, full.names=TRUE)
  pls.abs = list.files(path = abs.dir, pattern = glob2rx(paste0(sp,"*goingto.rda")), recursive=FALSE, full.names=TRUE)
  min.abs = list.files(path = abs.dir, pattern = glob2rx(paste0(sp,"*comingfrom.rda")), recursive=FALSE, full.names=TRUE)
  mfiles = list.files(path = migtime.dir, pattern = glob2rx(paste0(sp,"*migration*")), recursive=FALSE, full.names=TRUE)
  
  #import migration dates from previous analysis
  if(sp == "rthu"){ migdates = read.table(mfiles[1], header=TRUE, sep=",", quote="", fill=TRUE, as.is=TRUE, comment.char="")}
  if(sp %in% c("bchu", "bthu", "cahu", "ruhu")){ migdates = read.table(mfiles[2], header=TRUE, sep=",", quote="", fill=TRUE, as.is=TRUE, comment.char="")}
  migdates=cleanColNames(migdates)
  
  #load all the files into a list and remove rows with incomplete predictors
  abs = importANDformat(spfiles[1], 0, migdates, alpha_window)
  abs <- abs[which(complete.cases(subset(abs, select = -yday1))),]
  pres = importANDformat(spfiles[2], 1, migdates, alpha_window)
  pres <- pres[which(complete.cases(subset(pres, select = -yday1))),]
  min.abs = get(load(min.abs))
  min.abs = min.abs[which(complete.cases(min.abs)),]
  pls.abs = get(load(pls.abs))
  pls.abs = pls.abs[which(complete.cases(pls.abs)),]
  print (paste0("Imported data for species: ", sp))
  
  #subset data for models - note that models will be run forward and backwards within a season
  pa = subset(rbind(pres, abs), season %in% c("spring", "breeding", "fall"))
  pmin = subset(rbind(pres, min.abs), season %in% c("spring", "breeding", "fall"))
  ppls = subset(rbind(pres, pls.abs), season %in% c("spring", "breeding", "fall"))
  
  pa.spring = subset(rbind(pres, abs), season=="spring")
  pmin.spring = subset(rbind(pres, min.abs), season=="spring")
  ppls.spring = subset(rbind(pres, pls.abs), season=="spring")
  
  pa.fall = subset(rbind(pres, abs), season=="fall")
  pmin.fall = subset(rbind(pres, min.abs), season=="fall")
  ppls.fall = subset(rbind(pres, pls.abs), season=="fall")
  
  # presence data only to use for plots
  spr_fal = pres[pres$season=="spring" | pres$season=="fall",]
  
  # load statistical results from gamlss model evaluation (mod.list)
  # where: pmin=backwards comparison (all past points), ppls = forwards comparison (all future points)
  #        and, pa=local comparison (within an alpha hull)
  #load(paste0(stat.dir, "/", sp, ".mod.list.rda"))
```

# Data Summary
```{r, echo=FALSE}
# print(sp)
# print(paste0("Number occurrences: ", nrow(pres)))
# print(paste0("Number within-region absences from complete checklists: ", nrow(abs)))  
# print(paste0("Number absences for backward trajectory (past locations): ", nrow(min.abs))) 
# print(paste0("Number absences for forward trajectory (future locations): ", nrow(pls.abs))) 
#       
# print(paste0("Number spr/fall occurrences: ", nrow(spr_fal)))
# print(paste0("Number spr/fall within-region absences: ", nrow(abs[abs$season=="spring" | abs$season=="fall",]))) 
# print(paste0("Number spr/fall backward trajectory (past locations): ", nrow(min.abs[min.abs$season=="spring" | min.abs$season=="fall",]))) 
# print(paste0("Number spr/fall forward trajectory (future locations): ", nrow(pls.abs[pls.abs$season=="spring" | pls.abs$season=="fall",])))
```
Description                                   | value                 
--------------------------------------------  | ----------------
Number occurrences                            | `r nrow(pres)`
Number local background points from complete checklists | `r nrow(abs)`
Number background points for past occurrences (backward, min) | `r nrow(min.abs)`
Number background points for future occurrences (forward, pls) | `r nrow(pls.abs)`
Number of spring or autumn occurrences        | `r nrow(spr_fal)`
Number of spring or autumn local background points | `r nrow(abs[abs$season=="spring" | abs$season=="fall",])`
Number of spring or autumn backward points    | `r nrow(min.abs[min.abs$season=="spring" | min.abs$season=="fall",])`
Number of spring or autumn forward points    | `r nrow(min.abs[pls.abs$season=="spring" | pls.abs$season=="fall",])`


###Environment distribution by season
####EVI for occurrences
```{r, echo=FALSE}
sm=round(mean(spr_fal[spr_fal$season=="spring",]$EVI,),2)
ssd=round(sd(spr_fal[spr_fal$season=="spring",]$EVI,),2)
fm=round(mean(spr_fal[spr_fal$season=="fall",]$EVI,),2)
fsd=round(sd(spr_fal[spr_fal$season=="fall",]$EVI,),2)
ggplot(spr_fal, aes(EVI)) + geom_density(aes(fill=season), alpha=0.50) + ggtitle(paste0("spr (",sm,", ",ssd,") aut (",fm,", ",fsd,")")) + scale_fill_viridis(discrete=TRUE) + theme_bw()
print(paste0("The 95 percentiles are: ", quantile(spr_fal$EVI, 0.05, 0.95)))
```

####Temperature for occurrences
```{r, echo=FALSE}
sm=round(mean(spr_fal[spr_fal$season=="spring",]$t10m-273.15,),2)
ssd=round(sd(spr_fal[spr_fal$season=="spring",]$t10m-273.15,),2)
fm=round(mean(spr_fal[spr_fal$season=="fall",]$t10m-273.15,),2)
fsd=round(sd(spr_fal[spr_fal$season=="fall",]$t10m-273.15,),2)
ggplot(spr_fal, aes(t10m-273.15)) + geom_density(aes(fill=season), alpha=0.50) + ggtitle(paste0("spr (",sm,", ",ssd,") aut (",fm,", ",fsd,")")) + xlab("Celsius") + scale_fill_viridis(discrete=TRUE) + theme_bw()
print(paste0("The 95 percentiles are: ", quantile(spr_fal$t10m, 0.05, 0.95)))
```

####Elevation for occurrences
```{r, echo=FALSE}
sm=round(mean(spr_fal[spr_fal$season=="spring",]$SRTM_elev,),2)
ssd=round(sd(spr_fal[spr_fal$season=="spring",]$SRTM_elev,),2)
fm=round(mean(spr_fal[spr_fal$season=="fall",]$SRTM_elev,),2)
fsd=round(sd(spr_fal[spr_fal$season=="fall",]$SRTM_elev,),2)
ggplot(spr_fal, aes(SRTM_elev)) + geom_density(aes(fill=season), alpha=0.50) + ggtitle(paste0("spr (",sm,", ",ssd,") aut (",fm,", ",fsd,")")) + xlab("Elevation (m)") + scale_fill_viridis(discrete=TRUE) + theme_bw()
print(paste0("The 95 percentiles are: ", quantile(spr_fal$SRTM_elev, 0.05, 0.95)))
```

#Environment by day 
black points are environmental conditions at forward and backward comparison locations
```{r, echo=FALSE}
ggplot(pres, aes(yday, EVI)) + geom_point(data=min.abs, alpha=0.10) + geom_point(data=pls.abs, alpha=0.10) + 
  geom_point(alpha=0.10, aes(col=season)) + ggtitle(sp) +
  stat_smooth(data=pres, method="gam", formula = y~s(x, k=10), col="black") + xlab("Julian Day of the Year") +
  geom_vline(data=migdates, aes(xintercept=peak_lat)) + 
  scale_color_manual(values=c("#ffff99","#fdc086", "#7fc97f", "#beaed4")) +
  guides(color = guide_legend(override.aes= list(alpha = 1))) + theme_bw()

ggplot(pres, aes(yday, t10m-273.15)) + geom_point(data=min.abs, alpha=0.10) + 
  geom_point(data=pls.abs, alpha=0.10) +  xlab("Julian Day of the Year") +
  geom_point(alpha=0.10, aes(col=season)) + ggtitle(sp) + ylab("Temperature at 10 m (C)") +
  stat_smooth(data=pres, method="gam", formula = y~s(x, k=10), col="black") + 
  geom_vline(data=migdates, aes(xintercept=peak_lat)) + 
  scale_color_manual(values=c("#ffff99","#fdc086", "#7fc97f", "#beaed4")) +
  guides(color = guide_legend(override.aes= list(alpha = 1))) + theme_bw()

ggplot(pres, aes(yday, SRTM_elev)) + geom_point(data=min.abs, alpha=0.10) + geom_point(data=pls.abs, alpha=0.10) + 
  geom_point(alpha=0.10, aes(col=season)) + ggtitle(sp) + ylab("Elevation (m)") + xlab("Julian Day of the Year") +
  stat_smooth(data=pres, method="gam", formula = y~s(x, k=10), col="black") + 
  geom_vline(data=migdates, aes(xintercept=peak_lat)) + 
  scale_color_manual(values=c("#ffff99","#fdc086", "#7fc97f", "#beaed4")) +
  guides(color = guide_legend(override.aes= list(alpha = 1))) + theme_bw()
```

# Correlation of environmental variables, for Spring and Fall together
Note: This is all the data, presence, nearby, backward, and forward

```{r, echo=FALSE}
spr_fal_abs=abs[abs$season=="spring" | abs$season=="fall",]
spr_fal_min=min.abs[min.abs$season=="spring" | min.abs$season=="fall",]
spr_fal_pls=pls.abs[pls.abs$season=="spring" | pls.abs$season=="fall",]
spr_fal_cor = rbind(spr_fal_abs, spr_fal_min, spr_fal_pls)
spr_fal_cor$season="X"
spr_fal_cor = rbind(spr_fal_cor, spr_fal)

cor=cor(scale(spr_fal_cor$SRTM_elev), scale(spr_fal_cor$t10m))
ggplot(spr_fal_cor, aes(as.vector(scale(t10m)), as.vector(scale(SRTM_elev)))) + geom_point(alpha=0.10, aes(col=season)) + ggtitle(paste0("Pearson Correlation = ", round(cor,2))) + 
  xlab("scaled Temperature") + ylab("scaled Elevation") + 
  scale_color_manual(values=c("#fdc086", "#7fc97f", "black")) +
  guides(color = guide_legend(override.aes= list(alpha = 1))) + theme_bw()

cor=cor(scale(spr_fal_cor$SRTM_elev), scale(spr_fal_cor$EVI))
ggplot(spr_fal_cor, aes(as.vector(scale(EVI)), as.vector(scale(SRTM_elev)))) + geom_point(alpha=0.10, aes(col=season)) + ggtitle(paste0("Pearson Correlation = ", round(cor,2))) + 
  xlab("scaled EVI") + ylab("scaled Elevation") +  
  scale_color_manual(values=c("#fdc086", "#7fc97f", "black")) +
  guides(color = guide_legend(override.aes= list(alpha = 1))) + theme_bw()

cor=cor(scale(spr_fal_cor$t10m), scale(spr_fal_cor$EVI))
ggplot(spr_fal_cor, aes(as.vector(scale(EVI)), as.vector(scale(t10m)))) + geom_point(alpha=0.10, aes(col=season)) + ggtitle(paste0("Pearson Correlation = ", round(cor,2))) + 
  xlab("scaled EVI") + ylab("scaled Temperature") +  
  scale_color_manual(values=c("#fdc086", "#7fc97f", "black")) +
  guides(color = guide_legend(override.aes= list(alpha = 1))) + theme_bw()

#print(dim(spr_fal_cor))
```

#Distribution of Environmental data
Note: This is all the data, presence, nearby, backward, and forward

Environment   | mean            |   sd                     |   5%ile                  |   95%ile            
------------- | ----------------| -------------------------| -------------------------| -------------------------
EVI           | `r round(mean(spr_fal_cor$EVI),2)` | `r sd(spr_fal_cor$EVI,2)` | `r round(quantile(spr_fal_cor$EVI, 0.05)[[1]], 2)` | `r round(quantile(spr_fal_cor$EVI, 0.95)[[1]], 2)`
Temperature   | `r round(mean(spr_fal_cor$t10m),2)` | `r sd(spr_fal_cor$t10m,2)` | `r round(quantile(spr_fal_cor$t10m, 0.05)[[1]], 2)` | `r round(quantile(spr_fal_cor$t10m, 0.95)[[1]], 2)`
Elevation     | `r round(mean(spr_fal_cor$SRTM_elev),2)` | `r sd(spr_fal_cor$SRTM_elev,2)` | `r round(quantile(spr_fal_cor$SRTM_elev, 0.05)[[1]], 2)` | `r round(quantile(spr_fal_cor$SRTM_elev, 0.95)[[1]], 2)`


# distribution of environmental data by season, separately
Note: This is all the data, presence, nearby, backward, and forward
```{r, echo=FALSE}

spr_all = rbind(subset(pres, season=="spring"), subset(abs, season=="spring"), subset(min.abs, season=="spring"), subset(pls.abs, season=="spring"))

fall_all = rbind(subset(pres, season=="fall"), subset(abs, season=="fall"), subset(min.abs, season=="fall"), subset(pls.abs, season=="fall"))
```

Environment   | value   |   Spring                      | Autumn
------------- | --------| ----------------------------- |---------------------
EVI           | mean    |`r round(mean(spr_all$EVI, na.rm=TRUE),2)` | `r round(mean(fall_all$EVI),2)`
EVI           | sd      |`r round(sd(spr_all$EVI),2)`         | `r round(sd(fall_all$EVI),2)`
EVI           | 5 %ile  |`r round(quantile(spr_all$EVI, 0.05)[[1]],2)`|`r round(quantile(fall_all$EVI, 0.05)[[1]],2)`
EVI           | 95%ile  |`r round(quantile(spr_all$EVI, 0.95)[[1]],2)`|`r round(quantile(fall_all$EVI, 0.95)[[1]],2)`
Temperature   | mean    |`r round(mean(spr_all$t10m, na.rm=TRUE),2)` | `r round(mean(fall_all$t10m),2)`
Temperature   | sd      |`r round(sd(spr_all$t10m),2)`         | `r round(sd(fall_all$t10m),2)`
Temperature   | 5 %ile  |`r round(quantile(spr_all$t10m, 0.05)[[1]],2)`|`r round(quantile(fall_all$t10m, 0.05)[[1]],2)`
Temperature   | 95%ile  |`r round(quantile(spr_all$EVI, 0.95)[[1]],2)`|`r round(quantile(fall_all$EVI, 0.95)[[1]],2)`
Elevation     | mean    |`r round(mean(spr_all$SRTM_elev, na.rm=TRUE),2)` | `r round(mean(fall_all$SRTM_elev),2)`
Elevation     | sd      |`r round(sd(spr_all$SRTM_elev),2)`         | `r round(sd(fall_all$SRTM_elev),2)`
Elevation     | 5 %ile  |`r round(quantile(spr_all$SRTM_elev, 0.05)[[1]],2)`|`r round(quantile(fall_all$SRTM_elev, 0.05)[[1]],2)`
Elevation     | 95%ile  |`r round(quantile(spr_all$SRTM_elev, 0.95)[[1]],2)`|`r round(quantile(fall_all$SRTM_elev, 0.95)[[1]],2)`


# Map convex hulls of 5-day periods for Spring across years
```{r, echo=FALSE, warning=FALSE}
if(sp == "rthu"){myLocation = c(-105, 30, -60, 50)}
if(sp %in% c("bchu", "bthu", "cahu", "ruhu")){ myLocation = c(-130, 30, -105, 50)}

sp_hull = spr_fal[spr_fal$season=="spring",]

myLocation = c(-130, 30, -103, 50)
myMap = get_map(location=myLocation, source="stamen", maptype="toner", crop=FALSE)

chulls <- ddply(sp_hull, .(yday1), function(df) df[chull(df$location.long, df$location.lat), ])
ggmap(myMap) + geom_polygon(data=chulls, aes(x=location.long, y=location.lat, group=yday1, fill=month), col="black", alpha=0.25) + facet_wrap(~year, nrow=1) +
  scale_colour_brewer(palette = "Blues") + 
  scale_fill_brewer(palette = "Blues") + ggtitle (sp) + theme_map() + theme(legend.position ="bottom") +
  guides(color = guide_legend(override.aes= list(alpha = 1)))
```

# Map convex hulls of 5-day periods for Autumn across years
```{r, echo=FALSE, warning=FALSE}
if(sp == "rthu"){myLocation = c(-105, 30, -60, 50)}
if(sp %in% c("bchu", "bthu", "cahu", "ruhu")){ myLocation = c(-130, 30, -105, 50)}

sp_hull = spr_fal[spr_fal$season=="fall",]

myLocation = c(-130, 30, -103, 50)
myMap = get_map(location=myLocation, source="stamen", maptype="toner", crop=FALSE)

chulls <- ddply(sp_hull, .(yday1), function(df) df[chull(df$location.long, df$location.lat), ])
ggmap(myMap) + geom_polygon(data=chulls, aes(x=location.long, y=location.lat, group=yday1, fill=month), col="black", alpha=0.25) + facet_wrap(~year, nrow=1) +
  scale_colour_brewer(palette = "Oranges") + 
  scale_fill_brewer(palette = "Oranges") + ggtitle(sp) + theme_map() + theme(legend.position ="bottom") +
  guides(fill = guide_legend(override.aes= list(alpha = 1)))
```


example of within-area presence, in the middle of 2014 spring migration (February 24-28)
```{r, echo=FALSE}
# if(sp == "rthu"){myLocation = c(-105, 30, -60, 50)}
# if(sp %in% c("bchu", "bthu", "cahu", "ruhu")){ myLocation = c(-130, 30, -105, 50)}
# 
# myMap = get_map(location=myLocation, source="stamen", maptype="watercolor", crop=FALSE)
# 
# spr_example = pa.spring[pa.spring$year==2014 & pa.spring$window==8,]
# ggmap(myMap) + geom_point(data=spr_example, aes(x=location.long, y=location.lat, col=pres))
# 
# spr_mig_b = pmin.spring[pmin.spring$year==2014 & pmin.spring$window==8,]
# ggmap(myMap) + geom_point(data=spr_mig_b, aes(x=location.long, y=location.lat, col=pres))
# 
# spr_mig_f = ppls.spring[ppls.spring$year==2014 & ppls.spring$window==8,]
# ggmap(myMap) + geom_point(data=spr_mig_f, aes(x=location.long, y=location.lat, col=SRTM_elev))
```

Example of an animated map, for 2014, colored by temperature
```{r, echo=FALSE}
# 
# spr_fal_14 = spr_fal[spr_fal$year == 2014,]
# myMap = get_map(location=myLocation, source="stamen", maptype="watercolor", crop=FALSE)
# 
# mp = ggmap(myMap) + geom_point(data=spr_fal, aes(x=location.long, y=location.lat, frame=window, col=t10m-273.15)) + 
#   scale_color_gradient(low="blue", high="firebrick") 
# #gg_animate(mp)
```



