---
title: "hb-data-vis-allspecies.Rmd"
author: "Sarah Supp"
date: "April 25, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

## Getting set up:
Data Note: pmin is backwards (all previous points) and ppls is forwards (all future points)
```{r, echo=FALSE}
library(dplyr)
library(ggplot2)
library(ggmap)
library(gganimate)
library(lubridate)
library(plyr)
library(plotly)
library(gridExtra)
library(viridis)
library(ggthemes)
dropbox <- "/home/sarah/Dropbox/Hummingbirds/"
github <- "/home/sarah/Documents/GitHub/hb-migration/"

#assign based on the time frame (number of days) used to compute the alpha hulls
alpha_window = 5

# define pathnames (FAL will need to change pathnames)
function.dir <- paste0(github, "hb_RS_functions.R")
agan.dir <- paste0(dropbox, "hb_migration_data/ebird_annotated_raw/combined/")
abs.dir <- paste0(dropbox, "hb_migration_data/pseudoabsences/t", alpha_window, "/annotated/")
migtime.dir <- paste0(dropbox, "hb_migration_data/ebird_raw/eBird_checklists_2008-2014/aggregate_by_species/")
fig.dir <- paste0(dropbox, "NASA_Hummingbirds/P10_eBird_Migration_multiple topics/2-Mechanisms/figures/")
stat.dir <- paste0(dropbox, "NASA_Hummingbirds/P10_eBird_Migration_multiple topics/2-Mechanisms/stat-tests/")

source(function.dir) 

#list of species codes
species <- c("bchu", "bthu", "cahu", "ruhu", "rthu")
```


#Figure 1a. 
Map all hummingbirds in all years
```{r}
species <- c("bchu", "bthu", "cahu", "ruhu", "rthu")
all = NA
allabs = NA
allmin = NA
allpls = NA

for (sp in species){ 
  spfiles = list.files(path = agan.dir, pattern = glob2rx(paste0(sp,"*.RData")), recursive=FALSE, full.names=TRUE)
  mfiles = list.files(path = migtime.dir, pattern = glob2rx(paste0(sp,"*migration*")), recursive=FALSE, full.names=TRUE)
  pls.abs = list.files(path = abs.dir, pattern = glob2rx(paste0(sp,"*goingto.rda")), recursive=FALSE, full.names=TRUE)
  min.abs = list.files(path = abs.dir, pattern = glob2rx(paste0(sp,"*comingfrom.rda")), recursive=FALSE, full.names=TRUE)
 
  if(sp == "rthu"){ 
    migdates = read.table(mfiles[1], header=TRUE, sep=",", quote="", 
                         fill=TRUE, as.is=TRUE, comment.char="")}
  if(sp %in% c("bchu", "bthu", "cahu", "ruhu")){ 
     migdates = read.table(mfiles[2], header=TRUE, sep=",", quote="", 
                         fill=TRUE, as.is=TRUE, comment.char="")}
  migdates=cleanColNames(migdates)
   #load all the files into a list and remove rows with incomplete predictors
  pres = importANDformat(spfiles[2], 1, migdates, alpha_window)
  pres <- pres[which(complete.cases(subset(pres, select = -yday1))),]
  abs = importANDformat(spfiles[1], 0, migdates, alpha_window)
  abs <- abs[which(complete.cases(subset(abs, select = -yday1))),]
  min.abs = get(load(min.abs))
  min.abs = min.abs[which(complete.cases(min.abs)),]
  pls.abs = get(load(pls.abs))
  pls.abs = pls.abs[which(complete.cases(pls.abs)),]
  
  pres$sp = sp
  abs$sp = sp
  min.abs$sp = sp
  pls.abs$sp = sp
  
  all = rbind(all, pres)
  allabs = rbind(allabs, abs)
  allmin = rbind(allmin, min.abs)
  allpls = rbind(allpls, pls.abs)
}
all = all[-1,]
allabs = allabs[-1,]
allmin = allmin[-1,]
allpls = allpls[-1,]

all_spring = all[all$season=="spring",]
all_fall = all[all$season=="fall",]
all_breed = all[all$season=="breeding",]

myLocation = c(-130, 30, -60, 50)
myMap = get_map(location=myLocation, source="stamen", maptype="toner", crop=FALSE)

springmap = ggmap(myMap) + geom_point(data=all_spring, aes(x=location.long, y=location.lat, col=sp), alpha=0.25) + 
  scale_color_viridis(discrete=TRUE, option="B") +
  guides(color = guide_legend(override.aes= list(alpha = 1))) + theme_map() + theme(legend.position="none")
springmap

fallmap = ggmap(myMap) + geom_point(data=all_fall, aes(x=location.long, y=location.lat, col=sp), alpha=0.25) + 
  scale_color_viridis(discrete=TRUE, option="B") +
  guides(color = guide_legend(override.aes= list(alpha = 1))) + theme_map() + theme(legend.position="none")
fallmap

breedmap = ggmap(myMap) + geom_point(data=all_breed, aes(x=location.long, y=location.lat, col=sp)) +
  scale_color_viridis(discrete=TRUE, option="B" )+
  guides(color = guide_legend(override.aes= list(alpha = 1)))  +
  theme_map() + theme(legend.position ="bottom") 
breedmap
```

#Figure 1c. 2D version. 
Plot environmental values for an example window and its absence comparison
EVI and t10m as main axies, colored by elevation. Uses color-blind friendly shading schema
```{r}
ruhu_2013_94<- all %>%
  filter(sp=="ruhu", year==2013, yday1==94)

ruhu_2013_94_abs <- allabs %>%
  filter(sp=="ruhu", year==2013, yday1==94)

ruhu_2013_94_min <- allmin %>%
  filter(sp=="ruhu", year==2013, yday==94)

ruhu_2013_94_pls <- allpls %>%
  filter(sp=="ruhu", year==2013, yday1==94)

ruhu_2013_94$pres <- "occurrence"
ruhu_2013_94_abs$pres <- "background"
ruhu_2013_94_min$pres <- "previous occurrence"
ruhu_2013_94_pls$pres <- "future occurrence"
exampledat <- rbind(ruhu_2013_94, ruhu_2013_94_abs, ruhu_2013_94_min, ruhu_2013_94_pls)

fig1panels <- ggplot(exampledat, aes(EVI, t10m)) + geom_point(aes(color=SRTM_elev), size=3, alpha=0.5) + 
  scale_x_continuous(limits=c(0,0.7)) + scale_y_continuous(limits=c(265, 300)) + scale_color_viridis() + 
  theme_bw() + ylab("Temperature") +
  facet_wrap(~pres)

fig1panels
```

#Figure 1c. 3D version. 
Plot environmental values for an example window and its absence comparison
EVI, t10m, and Elevation as main axies, colored by latitude. Uses color-blind friendly shading schema
```{r}

p <- plot_ly(ruhu_2013_94, x = ~EVI, y = ~t10m, z = ~SRTM_elev, color=~location.lat) %>%
  add_markers() %>%
  layout(scene = list(xaxis = list(title = 'EVI'),
                     yaxis = list(title = 'Temperature'),
                     zaxis = list(title = 'Elevation')))

pabs <- plot_ly(ruhu_2013_94_abs, x = ~EVI, y = ~t10m, z = ~SRTM_elev, color=~location.lat) %>%
  add_markers() %>%
  layout(scene = list(xaxis = list(title = 'EVI'),
                     yaxis = list(title = 'Temperature'),
                     zaxis = list(title = 'Elevation')))

pmin <- plot_ly(ruhu_2013_94_min, x = ~EVI, y = ~t10m, z = ~SRTM_elev, color=~location.lat) %>%
  add_markers() %>%
  layout(scene = list(xaxis = list(title = 'EVI'),
                     yaxis = list(title = 'Temperature'),
                     zaxis = list(title = 'Elevation')))

ppls <- plot_ly(ruhu_2013_94_pls, x = ~EVI, y = ~t10m, z = ~SRTM_elev, color=~location.lat) %>%
  add_markers() %>%
  layout(scene = list(xaxis = list(title = 'EVI'),
                     yaxis = list(title = 'Temperature'),
                     zaxis = list(title = 'Elevation')))

#grid.arrange(p, pabs, pmin, ppls, nrow=1)
```



#Supplementary Figure: 
Plot difference of environment from mean occurrence values, for ALL species
```{r}

all_means <- all %>%
  filter(season %in% c("spring", "fall")) %>%
  select(year, yday1, season, EVI, SRTM_elev, t10m, sp) %>%
  group_by(sp, year, season, yday1) %>% 
  summarise_all(.funs = mean) 

all_sd <- all %>%
  filter(season %in% c("spring", "fall")) %>%
  select(year, yday1, season, EVI, SRTM_elev, t10m, sp) %>%
  group_by(sp, year, season, yday1) %>% 
  summarise_all(.funs = sd)

year = unique(all_means$year)
zdiffs = data.frame()

for (s in species){
  for (y in year) { 
    z_abs <- allabs %>%
      filter(season %in% c("spring", "fall")) %>%
      filter(sp == s & year == y) %>%
      select(yday1, EVI, SRTM_elev, t10m) 
      
    z_min <- allmin %>%
      filter(season %in% c("spring", "fall")) %>%
      filter(sp == s & year == y) %>%
      select(yday1, EVI, SRTM_elev, t10m) 
      
    z_pls <- allpls %>%
      filter(season %in% c("spring", "fall")) %>%
      filter(sp == s & year == y) %>%
      select(yday1, EVI, SRTM_elev, t10m) 
      
    zday = data.frame()
    
    for (d in unique(z_abs$yday1)) {
      # find the mean and sd of the occurrence data for the temporal window of species and year
      mean = all_means[all_means$sp==s & all_means$year==y & all_means$yday1==d, ]
      sd = all_sd[all_sd$sp==s & all_sd$year==y & all_sd$yday1==d, ]
      
      #in case there is a mismatch in yday1 between the pres and abs data
      if(nrow(mean) <1) {
        print(paste0("yday1 ", d, " is not in occurrence data for ", s, " ", y))
        next}
      
      #standardize (z-values) the data for that temporal window by the occurrence data
      za = z_abs %>%
        filter(yday1 == d) %>%
        mutate(zEVI = (EVI - mean$EVI)/sd$EVI,
               zSRTM = (SRTM_elev - mean$SRTM_elev)/sd$SRTM_elev,
               zt10m = (t10m - mean$t10m)/sd$t10m,
               type="local")
      
      zm = z_min %>%
        filter(yday1 == d) %>%
        mutate(zEVI = (EVI - mean$EVI)/sd$EVI,
               zSRTM = (SRTM_elev - mean$SRTM_elev)/sd$SRTM_elev,
               zt10m = (t10m - mean$t10m)/sd$t10m, 
               type="backward")
      
      zp = z_pls %>%
        filter(yday1 == d) %>%
        mutate(zEVI = (EVI - mean$EVI)/sd$EVI,
               zSRTM = (SRTM_elev - mean$SRTM_elev)/sd$SRTM_elev,
               zt10m = (t10m - mean$t10m)/sd$t10m,
               type="forward")
      
      zday = bind_rows(zday, za, zm, zp)
    }
  zday$sp = s
  zday$year = y
  
  zdiffs = bind_rows(zdiffs, zday)
  }
}

ggplot(zdiffs, aes(yday1, zEVI, group=interaction(year))) + stat_smooth(aes(fill=year, col=year)) + facet_wrap(~sp + type, ncol=3) + scale_y_continuous(limits=c(-2,2)) + geom_hline(yintercept=0) + theme(legend.position="bottom") 

ggplot(zdiffs, aes(yday1, zSRTM, group=interaction(year))) + stat_smooth(aes(fill=year, col=year)) + facet_wrap(~sp + type, ncol=3) + scale_y_continuous(limits=c(-2,2)) + geom_hline(yintercept=0) + theme(legend.position="bottom")

ggplot(zdiffs, aes(yday1, zt10m, group=interaction(year))) + stat_smooth(aes(fill=year, col=year)) + facet_wrap(~sp + type, ncol=3) + scale_y_continuous(limits=c(-2,2)) + geom_hline(yintercept=0) + theme(legend.position="bottom")
```


For fun: Animate all hummingbirds migrating in 2014
```{r}
all_14 = all[all$year==2014,]
myLocation = c(-130, 30, -60, 50)
myMap = get_map(location=myLocation, source="stamen", maptype="watercolor", crop=FALSE)

mp = ggmap(myMap) + geom_point(data=all_14, aes(x=location.long, y=location.lat, frame=month, col=sp), alpha=0.5)
gg_animate(mp)
```
