---
title: "hb-enviro-figures"
author: "Sarah Supp"
date: "7/9/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

## Getting set up:
Data Note: pmin is preceding (previous locations) and ppls is following (future locations)
```{r, echo=FALSE}
library(tidyverse)
library(gamlss)
library(ggmap)
library(gganimate) # devtools::install_github("dgrtwo/gganimate")
library(plyr)
library(plotly)
library(gridExtra)
library(maps)
library(mapdata)
library(viridis)
library(ggthemes)
library(cowplot)
```


##Get the data
...Add your pathnames to each downloaded R or data file
```{r}
#assign based on the time frame (number of days) used to compute the alpha hulls
alpha_window = 5

# define pathnames
function.dir <- paste0(pathname, "hb_RS_functions.R")
agan.dir <- paste0(pathname, "hb_migration_data/ebird_annotated_raw/combined/")
abs.dir <- paste0(pathname, "hb_migration_data/pseudoabsences/t", alpha_window, "/annotated/")
migtime.dir <- paste0(pathname, "hb_migration_data/ebird_raw/eBird_checklists_2008-2014/aggregate_by_species/")
fig.dir <- paste0(pathname, "NASA_Hummingbirds/P10_eBird_Migration_multiple topics/2-Mechanisms/figures/")
stat.dir <- paste0(pathname, "NASA_Hummingbirds/P10_eBird_Migration_multiple topics/2-Mechanisms/stat-tests/")

source(function.dir) 

#list of species codes
species <- c("bchu", "bthu", "cahu", "ruhu", "rthu")
```

## Import the environmental data for occurrence locations and for each of the three migration scenarios (immediate, preceding, following)
```{r}
all = NA
allabs = NA
allmin = NA
allpls = NA

for (sp in species){ 
  spfiles = list.files(path = agan.dir, pattern = glob2rx(paste0(sp,"*.RData")), recursive=FALSE, full.names=TRUE)
  mfiles = list.files(path = migtime.dir, pattern = glob2rx(paste0(sp,"*migration*")), recursive=FALSE, full.names=TRUE)
  pls.abs = list.files(path = abs.dir, pattern = glob2rx(paste0(sp,"*goingto.rda")), recursive=FALSE, full.names=TRUE)
  min.abs = list.files(path = abs.dir, pattern = glob2rx(paste0(sp,"*comingfrom.rda")), recursive=FALSE, full.names=TRUE)
 
  if(sp == "rthu"){ 
    migdates = read.table(mfiles[1], header=TRUE, sep=",", quote="", 
                         fill=TRUE, as.is=TRUE, comment.char="")}
  if(sp %in% c("bchu", "bthu", "cahu", "ruhu")){ 
     migdates = read.table(mfiles[2], header=TRUE, sep=",", quote="", 
                         fill=TRUE, as.is=TRUE, comment.char="")}
  migdates=cleanColNames(migdates)
  
   #load all the files into a list and remove rows with incomplete predictors
  pres = importANDformat(spfiles[2], 1, migdates, alpha_window)
  pres <- pres[which(complete.cases(subset(pres, select = -yday1))),]
  abs = importANDformat(spfiles[1], 0, migdates, alpha_window)
  abs <- abs[which(complete.cases(subset(abs, select = -yday1))),]
  min.abs = get(load(min.abs))
  min.abs = min.abs[which(complete.cases(min.abs)),]
  pls.abs = get(load(pls.abs))
  pls.abs = pls.abs[which(complete.cases(pls.abs)),]
  
  pres$sp = sp
  abs$sp = sp
  min.abs$sp = sp
  pls.abs$sp = sp
  
  all = rbind(all, pres)
  allabs = rbind(allabs, abs)
  allmin = rbind(allmin, min.abs)
  allpls = rbind(allpls, pls.abs)
}
all = all[-1,]
allabs = allabs[-1,]
allmin = allmin[-1,]
allpls = allpls[-1,]

all_spring = all[all$season=="spring",]
all_fall = all[all$season=="fall",]
all_breed = all[all$season=="breeding",]
```

# Summarize the results from the gamlss analysis (from hb-gamlss-eval.R)
Get the results table out for the best fitting gamlss models. 
```{r}
res_files <- list.files(stat.dir, pattern = "mod.list.rda", full.names = TRUE)
full_res <- ldply(res_files, function(x) {
  load(x)
  sp <- substr(gsub(stat.dir, "", x), 1, 4)
  out <- ldply(mod.list, function(y) {
    y$mod <- row.names(y)
    return(y)})
  out$species <- sp
  return(out)})
full_res <- mutate(full_res, data = .id) %>% 
  select(species, data, mod, AIC, w, Rsq, df, nobs)
write.csv(full_res, file=paste0(stat.dir, "full_results_table.csv"))
best_mods <- group_by(full_res, species, data) %>% slice(which.min(AIC))
```


#Figure 1. Data Figure
Map of hummingbird locations for each species in spring, breeding, and autumn seasons. Occurrences are shown for all years combined, although data was analyzed within each year separately.
```{r}
# change naming for the figure (e.g. fall should be autumn, scientific names)
map_all <- all[all$season %in% c("spring", "breeding", "fall"),]
map_all$season <- factor(map_all$season, levels = c("spring", "breeding", "fall", "autumn"))
map_all[map_all == "fall"] <- "autumn"
map_all$season <- factor(map_all$season, levels = c("spring", "breeding", "autumn"))
map_all[map_all == "bchu"] <- "A. alexandri"
map_all[map_all == "bthu"] <- "S. platycercus"
map_all[map_all == "cahu"] <- "S. calliope"
map_all[map_all == "ruhu"] <- "S. rufus"
map_all[map_all == "rthu"] <- "A. colubris"

#plot a map with simple borders
map = ggplot() + geom_polygon(data=map_data("world"), aes(x=long, y=lat, group=group), fill="gray") + coord_fixed(ratio=1.3, xlim=c(-130,-50), ylim=c(20,60)) + theme_map()

# Map occurrence locations in Black and White
map + geom_point(data=map_all, aes(x=location.long, y=location.lat), alpha=0.25, size=0.25) +
  theme_map() + theme(legend.position="none") + facet_grid(sp ~ season) +
  theme(strip.text = element_text(size = 12))
```


#Figure 3. 
Plot the parameter estimates for the predictor variables in the gamlss models, for ALL species
```{r}
res_files <- list.files(stat.dir, pattern="gamlss.list.rda", full.names = TRUE)

res <- ldply(res_files, function(x) {
  load(x)
  f <- gsub(stat.dir, "", x)
  dat <- substr(f, 7, nchar(f) - 16)
  sp <- substr(f, 1, 4)
  # this is needed for those where we saved any significant model (i.e. gamlss.list > 1)
  best_mod <- filter(best_mods, species == sp, data == dat)
  mod <- gamlss.list[[best_mod$mod]]
  out <- broom::tidy(mod)
  out$species <- sp
  out$dir.season <- dat
  return(out)
})

res <- separate(res, dir.season, into=c("direction", "season"), sep = "\\.")
res <- filter(res, term!="(Intercept)")
res$term <- gsub(":", "*", gsub("\\)", "", gsub("pb\\(", "", res$term)))
# we have EVI*SRTM_elev and SRTM_elev*EVI - these need to be the same. 
res$term <- ifelse(res$term=="SRTM_elev*EVI", "EVI*SRTM_elev", res$term)
res$term <- factor(res$term, levels=c("EVI", "t10m", "SRTM_elev", "EVI*t10m", "EVI*SRTM_elev", "t10m*SRTM_elev", "EVI*t10m*SRTM_elev"))

#fix labels to match text
res_preds <- res2[res2$term %in% c("EVI", "t10m", "SRTM_elev"),]
res_preds$term <- factor(res_preds$term, levels = c("EVI", "t10m", "SRTM_elev", "Temperature", "Elevation"))
res_preds[res_preds == "t10m"] <- "Temperature"
res_preds[res_preds == "SRTM_elev"] <- "Elevation"
res_preds$term <- factor(res_preds$term, levels = c("Temperature", "EVI", "Elevation"))
res_preds$term <- ordered(res_preds$term, levels = c("Temperature", "EVI", "Elevation"))

res_preds$direction <- factor(res_preds$direction, levels = c("local", "backward", "forward", "preceding", "immediate", "following"))
res_preds[res_preds == "local"] <- "immediate"
res_preds[res_preds == "backward"] <- "preceding"
res_preds[res_preds == "forward"] <- "following"
res_preds$direction <- factor(res_preds$direction, levels = c("preceding", "immediate", "following"))

res_preds[res_preds == "/bch"] <- "A. alexandri"
res_preds[res_preds == "/bth"] <- "S. platycercus"
res_preds[res_preds == "/cah"] <- "S. calliope"
res_preds[res_preds == "/ruh"] <- "S. rufus"
res_preds[res_preds == "/rth"] <- "A. colubris"

#plot
pd <- position_jitter(width=NULL, height=0.25)
# reverse axes for better readability?
main3 <- ggplot(res_preds, aes(x = estimate, y = term, colour = direction, shape = direction, group = direction)) + 
 # geom_errorbar(aes(ymin=estimate-(1.96*std.error), ymax=estimate+(1.96*std.error)), colour="black", width=.1, position=pd) +
  geom_point(position=pd, size=3) + 
  geom_vline(xintercept = 0, linetype="dashed", col="black") + 
  geom_hline(yintercept=seq(1.5, length(unique(res_preds$term))-0.5, 1), #adds horizontal gridlines
             lwd=0.25, colour="gray") +
    scale_color_grey() +
#  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_grid(species~season)
```

Figure S1. 
Plot the parameter estimates for the interactions in the gamlss models, for ALL species
```{r}
# Supplementary figure with interactions
#fix labels to match text
res_interaxn <- res2[res2$term %in% c("EVI*t10m", "t10m*SRTM_elev", "EVI*SRTM_elev", "EVI*t10m*SRTM_elev"),]
res_interaxn$term <- factor(res_interaxn$term, levels = c("EVI*t10m", "t10m*SRTM_elev", "EVI*SRTM_elev", "EVI*t10m*SRTM_elev", "EVI * Temperature", "Temperature * Elevation", "EVI * Elevation", "EVI * Temperature * Elevation"))
res_interaxn[res_interaxn == "EVI*t10m"] <- "EVI * Temperature"
res_interaxn[res_interaxn == "EVI*SRTM_elev"] <- "EVI * Elevation"
res_interaxn[res_interaxn == "t10m*SRTM_elev"] <- "Temperature * Elevation"
res_interaxn[res_interaxn == "EVI*t10m*SRTM_elev"] <- "EVI * Temperature * Elevation"
res_interaxn$term <- factor(res_interaxn$term, levels = c("EVI * Temperature", "Temperature * Elevation", "EVI * Elevation", "EVI * Temperature * Elevation"))
res_interaxn$term <- ordered(res_interaxn$term, levels = c("EVI * Temperature", "Temperature * Elevation", "EVI * Elevation", "EVI * Temperature * Elevation"))

res_interaxn$direction <- factor(res_interaxn$direction, levels = c("local", "backward", "forward", "preceding", "immediate", "following"))
res_interaxn[res_interaxn == "local"] <- "immediate"
res_interaxn[res_interaxn == "backward"] <- "preceding"
res_interaxn[res_interaxn == "forward"] <- "following"
res_interaxn$direction <- factor(res_interaxn$direction, levels = c("preceding", "immediate", "following"))

res_interaxn[res_interaxn == "/bch"] <- "A. alexandri"
res_interaxn[res_interaxn == "/bth"] <- "S. platycercus"
res_interaxn[res_interaxn == "/cah"] <- "S. calliope"
res_interaxn[res_interaxn == "/ruh"] <- "S. rufus"
res_interaxn[res_interaxn == "/rth"] <- "A. colubris"

#plot
pd <- position_jitter(width=NULL, height=0.25)
# reverse axes for better readability?
supmat4 <- ggplot(res_interaxn, aes(x = estimate, y = term, colour = direction, shape = direction, group = direction)) + 
 # geom_errorbar(aes(ymin=estimate-(1.96*std.error), ymax=estimate+(1.96*std.error)), colour="black", width=.1, position=pd) +
  geom_point(position=pd, size=3) + 
  geom_vline(xintercept = 0, linetype="dashed", col="black") + 
  geom_hline(yintercept=seq(1.5, length(unique(res_interaxn$term))-0.5, 1), #adds horizontal gridlines
             lwd=0.25, colour="gray") +
    scale_color_grey() +
#  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_grid(species~season)
```


